---
title: "shiny + shinyjs"
author: "Paul Viefers"
date: "17. September 2015"
output: 
    ioslides_presentation:
      highlight: espresso
      smaller: yes
      widescreen: yes
      keepmd: true
runtime: shiny
---

## The RStudio & Shiny Universe

  * Shiny initially: interactive charts ([example](http://shiny.rstudio.com/gallery/kmeans-example.html)).
  * Nowadays lets you build full-flegded UIs ([example](http://shiny.rstudio.com/gallery/bus-dashboard.html)).
  * Interactive websites are great:
    + those familiar with JS might miss some functionality.
    + those not familiar might not be aware that they miss it.
  * `shinyjs` adds some functionality commonly encountered on the web.

## Basic shiny: first layer of interactivity

```{r, echo=FALSE}
inputPanel(
  selectInput("n_breaks", label = "Number of bins:",
              choices = c(10, 20, 35, 50), selected = 20),
  
  sliderInput("bw_adjust", label = "Bandwidth adjustment:",
              min = 0.2, max = 2, value = 1, step = 0.2)
)

renderPlot({
  hist(faithful$eruptions, probability = TRUE, breaks = as.numeric(input$n_breaks),
       xlab = "Duration (minutes)", main = "Geyser eruption duration")
  
  dens <- density(faithful$eruptions, adjust = input$bw_adjust)
  lines(dens, col = "blue")
})
```

## shiny + shinyjs: a second layer of interactivity

```{r, echo=FALSE, message=FALSE}
library(shinyjs)
session_password <- "koelsch"
shinyApp(
    ui = shinyUI(fluidPage(
        useShinyjs(),
        div( 
            id = "login_page",
            titlePanel("Welcome"),
            br(),
            sidebarLayout(
                
                sidebarPanel(
                    h2("Login"),
                    p("Please enter user name and password."),
                    hidden(
                        div(
                            id = "login_error",
                            span("Invalid. Please check for typos and try again.", style = "color:red")
                        )
                    )
                ),
                
                mainPanel(
                    textInput("user", "User", ""),
                    textInput("password", "Password", ""),
                    actionButton("login", "Login", class = "btn-primary")
                )
            )
        ),
        
        hidden(
            div( id = "interactive_chart_m",
                 selectInput("n_breaks", label = "Number of bins:",
                             choices = c(10, 20, 35, 50), selected = 20),
                 
                 sliderInput("bw_adjust", label = "Bandwidth adjustment:",
                             min = 0.2, max = 2, value = 1, step = 0.2),
                 plotOutput("Markus"),
                 actionLink("logout_m", "Logout")
            )
        ),
        
        hidden(
            div( id = "interactive_chart_k",
                 # Give the page a title
                titlePanel("Telephones by region"),
    
                # Generate a row with a sidebar
                sidebarLayout(      
                  
                  # Define the sidebar with one input
                  sidebarPanel(
                    selectInput("region", "Region:", 
                                choices=colnames(WorldPhones)),
                    hr(),
                    helpText("Data from AT&T (1961) The World's Telephones.")
                  ),
                  
                  # Create a spot for the barplot
                  mainPanel(
                    plotOutput("Kyrill"),
                    actionLink("logout_k", "Logout")
                  )
            )
        )
    )
    )
    ),
    
    server = function(input, output, session) {
        
        output$Markus <- renderPlot({
            hist(faithful$eruptions, probability = TRUE, breaks = as.numeric(input$n_breaks),
                 xlab = "Duration (minutes)", main = "Geyser eruption duration")
            
            dens <- density(faithful$eruptions, adjust = input$bw_adjust)
            lines(dens, col = "blue")
        })
        
        output$Kyrill <- renderPlot({
            # Render a barplot
            barplot(WorldPhones[,input$region]*1000,
                    main=input$region,
                    ylab="Number of Telephones",
                    xlab="Year")
        })
        
        # When the Login button is clicked, check whether user name is in list
        observeEvent(input$login, {
            
            # User-experience stuff
            shinyjs::disable("login")
            
            # Check whether user name is correct
            # Fix me: test against a session-specific password here, not username
            user_ok <- input$password==session_password
            users_allowed <- input$user %in% c('Markus', 'Kyrill')
            
            # If credentials are valid push user into experiment
            if(user_ok & users_allowed){
                shinyjs::hide("login_page")
                if(input$user == 'Markus'){
                    shinyjs::show("interactive_chart_m")
                } 
                if(input$user == 'Kyrill'){
                    shinyjs::show("interactive_chart_k")
                }
            } else {
                # If credentials are invalid throw error and prompt user to try again
                shinyjs::reset("login_page")
                shinyjs::show("login_error")
                shinyjs::enable("login")
            }
        })
        
        observeEvent(input$logout_m, {
            shinyjs::reset("login_page")
            shinyjs::enable("login")
            shinyjs::hide("interactive_chart_m")
            shinyjs::show("login_page")
        })
        
        observeEvent(input$logout_k, {
            shinyjs::reset("login_page")
            shinyjs::enable("login")
            shinyjs::hide("interactive_chart_k")
            shinyjs::show("login_page")
        })
    },
    
    options = list(height = 300)
)
```

## Code

- The usual structure: `ui.R` with user interface and `server.R` with code for server-side handling.
- `ui.R`: A series of `div()`s that are uncovered/covered sequentially.
- `server.R`: a set of commands and 'listeners' that handle the R code to be executed and also the JS commands to cover and uncover.
- See example code for experiment.

## Basic JS functionality just used

- `hidden(tagId)`: Create a Shiny tag that is invisible when the Shiny app starts. The tag can be made visible later
- `show(tagId)`: show makes an element visible, hide makes an element invisible.
- `disable`/`enable(tagId)`: Enable or disable an input element. A disabled element is not usable and not clickable.
- `reset(tagId)`: Reset any input element back to its original value.

## Make some inputs mandatory

```{r, echo=FALSE, message=FALSE}
library(shinyjs)
session_password <- "koelsch"
fieldsMandatory <- c('user', 'password')

shinyApp(
    ui = shinyUI(fluidPage(
        useShinyjs(),
        div( 
            id = "login_page",
            titlePanel("Welcome"),
            br(),
            sidebarLayout(
                
                sidebarPanel(
                    h2("Login"),
                    p("Please enter user name and password."),
                    hidden(
                        div(
                            id = "login_error",
                            span("Invalid. Please check for typos and try again.", style = "color:red")
                        )
                    )
                ),
                
                mainPanel(
                    textInput("user", "User", ""),
                    textInput("password", "Password", ""),
                    actionButton("login", "Login", class = "btn-primary")
                )
            )
        ),
        
        hidden(
            div( id = "interactive_chart_m",
                 selectInput("n_breaks", label = "Number of bins:",
                             choices = c(10, 20, 35, 50), selected = 20),
                 
                 sliderInput("bw_adjust", label = "Bandwidth adjustment:",
                             min = 0.2, max = 2, value = 1, step = 0.2),
                 plotOutput("Markus")
            )
        ),
        
        hidden(
            div( id = "interactive_chart_k",
                 selectInput("n_breaks", label = "Number of bins:",
                             choices = c(10, 20, 35, 50), selected = 20),
                 
                 sliderInput("bw_adjust", label = "Bandwidth adjustment:",
                             min = 0.2, max = 2, value = 1, step = 0.2),
                 plotOutput("Kyrill")
            )
        )
    )
    ),
    
    server = function(input, output, session) {
        
        output$Markus <- renderPlot({
            hist(faithful$eruptions, probability = TRUE, breaks = as.numeric(input$n_breaks),
                 xlab = "Duration (minutes)", main = "Geyser eruption duration")
            
            dens <- density(faithful$eruptions, adjust = input$bw_adjust)
            lines(dens, col = "blue")
        })
        
        output$Kyrill <- renderPlot({
            # Render a barplot
            barplot(WorldPhones[,input$region]*1000,
                    main=input$region,
                    ylab="Number of Telephones",
                    xlab="Year")
        })
        
        observe({
            # check if all mandatory fields have a value
            mandatoryFilled <-
            vapply(fieldsMandatory,
                function(x) {
                    !is.null(input[[x]]) && input[[x]] != ""
                },
                logical(1)
            )
            mandatoryFilled <- all(mandatoryFilled)
  
            # enable/disable the login button
            shinyjs::toggleState(id = "login", condition = mandatoryFilled)
        })
        
        # When the Login button is clicked, check whether user name is in list
        observeEvent(input$login, {
            
            # User-experience stuff
            shinyjs::disable("login")
            
            # Check whether user name is correct
            # Fix me: test against a session-specific password here, not username
            user_ok <- input$password==session_password
            users_allowed <- input$user %in% c('Markus', 'Kyrill')
            
            # If credentials are valid push user into experiment
            if(user_ok & users_allowed){
                shinyjs::hide("login_page")
                if(input$user == 'Markus'){
                    shinyjs::show("interactive_chart_m")
                } 
                if(input$user == 'Kyrill'){
                    shinyjs::show("interactive_chart_k")
                }
            } else {
                # If credentials are invalid throw error and prompt user to try again
                shinyjs::reset("login_page")
                shinyjs::show("login_error")
                shinyjs::enable("login")
            }
        })
    },
    
    options = list(height = 300)
)
```

## Other selected JS functions available

- `toggleState`: The element will be enabled if the `condition` evalutes to `TRUE` and disabled otherwise.

- `delay`: execute R code (including any shinyjs functions) after a specified amount of time.

- `info`: show a message to the user (using JavaScript's alert under the hood).

- `onclick`: run R code when an element is clicked. Was originally developed with the sole purpose of running a shinyjs function when an element is clicked, though any R code can be used.

- `onevent`: similar to onclick, but can be used with many other events instead of click.

- `logjs`: print a message to the JavaScript console (mainly used for debugging purposes).

- `extendShinyjs`: allows you to write your own JavaScript functions and use shinyjs to call them as if they were regular R code. More information is available in the section "Calling your own JavaScript functions from R".


## Wrap up

- `shinyjs` adds very useful functionality. 
- It's good to be aware of it.
- Very handy together with reactive expressions (`observe`, `observeEvent` et al.)
- Can code small surveys, simply end-user apps or experiments with little effort.
